Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-10-25 algorithm.tex
AA

It is not recommended to chose $\beta=1$, because it is possible that the
$k-1$-th and $k$-th Pareto front approximations are not very different but next ones could be. In other words,
it is unsafe to terminate as soon as the Pareto front does not considerably change
between two consecutive eras, thus a strategy taking into
account the history of the Pareto fronts has been adopted in the
proposed approach.
%!!! IN INGLESE E' PIU' CORRETTO DIRE: ``IT IS POSSIBLE THAT THE PARETO
%FRONTS WERE NOT VERY DIFFERENT''? (CON ``WERE'' AL POSTO DI ``ARE'')

\subsection{Exploration vs exploitation in PS}
Our algorithm is able to inherently balance exploration and exploitation, which are ``two cornerstones of problem solving by search''~\cite{eiben1998evolutionary}. While a detailed survey on the exploration vs exploitation trade-off can be found in~\cite{vcrepinvsek2013exploration}, here we limitedly recall that \emph{exploration} is the capability of an algorithm to consider different parts of the configuration space, not being confined only to a limited subset and being trapped in local optima. Following the terminology in \cite{weise2012evolutionary}, if the exploration is not sufficiently wide, we sat that the algorithm suffers from \emph{premature convergence}, i.e. it starts to analyze only a narrow portion of the configuration space, that may lead to ignore potential Pareto configurations that are outside this portion. When premature convergence happens, there is insufficient \emph{diversity}~\cite{weise2012evolutionary}, i.e. all the provided solutions are similar.
On the other hand, \emph{exploitation} is the capability to recognize and densely analyze the parts of configuration space that have many Pareto configurations.

As in genetic algorithms exploration and exploration are guaranteed by mutation, crossover and selection, in our algorithm these guarantees are provided for the following reasons:
	\begin{itemize}
		\item Merging and splitting operation are dynamic processes. A portion
of the parameter space may be populated by a large number of small
regions in some eras. However, after having been sufficiently explored, the innovation score of these regions may decrease and therefore that portion may be covered by a small number of large regions in next eras. This is a desirable property of PS, since, after a number of eras of intense exploration of a portion of configuration space, this exploration may turn to be enough thorough and so it may be time for other portions to be minutely scanned. Otherwise stated, PS dynamics can be summarized as follows: while eras alternate, PS zooms out from the configurations subspaces already sufficiently scanned in the past, whereas it zooms in subspaces which provide novelty (guaranteeing exploitation), meaning that they have not been properly analyzed yet (this helps exploration). 
		\item No portion of parameter space is neglected since, at each era, all the regions, even the least interesting ones, have the chance to be explored with new evaluated points (providing exploration), even though some regions are less densely explored than others.
	\end{itemize}

To summarize,thanks to the above characteristics, PS is able to recognize the regions when the best configurations are while not being trapped in only a narrow subset of configuration space.

A pseudo-code representation of PS algorithm is in Alg.\ref{alg:algo}.

@@

			\nl \\ // Splitting and merging operations\\
			\nl $\psi(\mathscr{R}_{i,h}) \Leftarrow \lbrace 
					\psi(R) | R\in \mathscr{R}_{i,h}	
				\rbrace$; \\
			\nl construct $\mathscr{M}_i$; // the set of pair of contiguous no innovation regions \\
			\nl $\mu(\mathscr{M}_i) = \bigcup_{(R_1,R_2)\in\mathscr{M}_i} \mu(R_1,R_2)$;\\
			\nl\\ // Prepare the configuration space partition for the next iteration \\
			\nl $\mathcal{R}_{i+1} \triangleq
	\psi\left(\mathscr{R}_{i,h}\right) \cup \mathcal{R}_{i,l}\cup
	\left(\mathcal{R}_{i,n}\setminus\bigcup_{\left(R_{1},R_{2}\right)\in\mathcal{C}}\left\{ R_{1},R_{2}\right\} \right)\cup \mu(\mathscr{M}_i)$;\\
		}
	\nl $\tilde{\mathscr{P}} \Leftarrow \mathscr{P}_i$
\end{algorithm}
%

@@

Another possible approach is closer to the ``Island scheme'' for GA parallelization~\cite{Borkar14parallel}. The configuration space is initially divided in $n$ macro-regions, each one assigned to a different processing node, which autonomously performs all the PS steps, including merging and splitting, considering the assigned macro-region as the entire parameter space. Note that, while the Island scheme for GA parallelization requires a migration operator, that is not easy to properly tune, we do not have special requirements. On the other hand, one of the macro-regions may begin to provide very low innovation at a certain era. In this case, dedicating to it the same number $K$ of test function as the other macro-regions is inefficient. To solve this problem, the number of test functions assigned to each macro-region should be perdiodically renegotiated, in favor of the macro-regions providing most innovation.

>>>>

@@

			\nl \\ // Splitting and merging operations\\
			\nl $\psi(\mathscr{R}_{i,h}) \Leftarrow \lbrace 
					\psi(R) | R\in \mathscr{R}_{i,h}	
				\rbrace$; \\
			\nl construct $\mathscr{M}_i$; // the set of pair of contiguous no innovation regions \\
			\nl $\mu(\mathscr{M}_i) = \bigcup_{(R_1,R_2)\in\mathscr{M}_i} \mu(R_1,R_2)$;\\
			\nl\\ // Prepare the configuration space partition for the next iteration \\
			\nl $\mathcal{R}_{i+1} \triangleq
	\psi\left(\mathscr{R}_{i,h}\right) \cup \mathcal{R}_{i,l}\cup
	\left(\mathcal{R}_{i,n}\setminus\bigcup_{\left(R_{1},R_{2}\right)\in\mathcal{C}}\left\{ R_{1},R_{2}\right\} \right)\cup \mu(\mathscr{M}_i)$;\\
		}
	\nl $\tilde{\mathscr{P}} \Leftarrow \mathscr{P}_i$
\end{algorithm}

It is not recommended to chose $\beta=1$, because it is possible that the
$k-1$-th and $k$-th Pareto front approximations are not very different but next ones could be. In other words,
it is unsafe to terminate as soon as the Pareto front does not considerably change
between two consecutive eras, thus a strategy taking into
account the history of the Pareto fronts has been adopted in the
proposed approach.

A pseudo-code representation of PS algorithm is in Alg.\ref{alg:algo}.
%!!! IN INGLESE E' PIU' CORRETTO DIRE: ``IT IS POSSIBLE THAT THE PARETO
%FRONTS WERE NOT VERY DIFFERENT''? (CON ``WERE'' AL POSTO DI ``ARE'')

\subsection{Exploration vs exploitation in PS}
Our algorithm is able to inherently balance exploration and exploitation, which are ``two cornerstones of problem solving by search''~\cite{eiben1998evolutionary}. While a detailed survey on the exploration vs exploitation trade-off can be found in~\cite{vcrepinvsek2013exploration}, here we limitedly recall that \emph{exploration} is the capability of an algorithm to consider different parts of the configuration space, not being confined only to a limited subset and being trapped in local optima. Following the terminology of \cite{weise2012evolutionary}, if the exploration is not sufficiently wide, the algorithm suffers from \emph{premature convergence}, i.e. it starts to analyze only a narrow portion of the configuration space, thus ignoring potential Pareto configurations that are outside this portion. When premature convergence happens, there is insufficient \emph{diversity}~\cite{weise2012evolutionary}, i.e. all the provided solutions are similar.
On the other hand, \emph{exploitation} is the capability to recognize and densely analyze the parts of configuration space that have many Pareto configurations.

While in genetic algorithms exploration and exploration are guaranteed by mutation, crossover and selection, in our algorithm these guarantees are provided for the following reasons:
	\begin{itemize}
		\item Merging and splitting operation are dynamic processes. A portion
of the parameter space may be populated by a large number of small
regions in some eras, and thus a large number of test configurations. However, after having been sufficiently explored, the innovation score of these regions may decrease, they may be merged and therefore that portion of configuration space may be covered by a small number of large regions in next eras. This is a desirable property of PS, since, after a number of eras of intense exploration of a configuration space part, this exploration may turn to be enough thorough and so it may be time for other parts to be minutely scanned. 
PS dynamics can be summarized as follows: while eras alternate, PS zooms out from the configurations subspaces already sufficiently scanned in the past, whereas it zooms in on subspaces which provide novelty (guaranteeing exploitation), meaning that they have not been properly analyzed yet (this helps exploration). 
		\item No portion of parameter space is neglected since, at each era, all the regions, even the least interesting ones, have the chance to be explored with new evaluated points (providing exploration), even though some regions are less densely explored than others.
	\end{itemize}

To summarize,thanks to the above characteristics, PS is able to recognize the regions when the best configurations are, while avoiding premature convergence, not being trapped in a narrow subset of configuration space.

%

@@

Another possible approach is closer to the ``Island scheme'' for GA parallelization~\cite{Borkar14parallel}. The configuration space is initially divided in $n$ macro-regions, each one assigned to a different processing node, which autonomously performs all the PS steps, including merging and splitting, considering the assigned macro-region as the entire parameter space. Note that, while the Island scheme for GA parallelization requires a migra0tion operator, that is not easy to properly tune, we do not have special requirements. On the other hand, one of the macro-regions may begin to provide very low innovation at a certain era. In this case, dedicating to it the same number $K$ of test function as the other macro-regions is inefficient. To solve this problem, the number of test functions assigned to each macro-region should be perdiodically renegotiated, in favor of the macro-regions providing most innovation.

### 2014-10-25 experiments.tex
AA

\subsection{Evaluation metrics}
In order to evaluate the Paretos from a quantitative point-of-view we
will consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}. Both of them permit to quantify the extent of the Pareto front that is an important factor to consider when evaluating the quality of a Pareto front. Indeed, authors in \cite{zitzler_ec00,weise2012evolutionary} carefully observe that an exploration algorithm should preferably provide Pareto fronts whose objective function range is large. Following the terminology of~\cite{weise2012evolutionary}, a \emph{uniform convergence} should be guaranteed.
Such Pareto fronts permit to investigate in a broader way the potentials of the designed device. This is important, in general, since if only a narrow portion of the Pareto front is provided due to non-uniform convergence, the decision maker may neglect some important options~\cite{weise2012evolutionary}. Having a narrow Pareto front becomes essential, in particular, when designing a device that may be required to work in different operational conditions, for example with stringent execution time constraints and less strict power dissipation ones or, on the contrary, with low power requirements and tolerance to high execution time.

>>>>

\subsection{Evaluation metrics}
In order to evaluate the Paretos from a quantitative point-of-view we
will consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}. Both of them permit to quantify the extent of the Pareto front that is an important factor to consider when evaluating the quality of a Pareto front. Indeed, authors in \cite{zitzler_ec00,weise2012evolutionary} observe that an exploration algorithm should preferably provide Pareto fronts whose objective function range is large. In other words, following the terminology of~\cite{weise2012evolutionary}, a \emph{uniform convergence} should be guaranteed.
Pareto fronts with large objective function ranges permit to investigate in a broader way the potentials of the designed device. This is important, in general, since if only a narrow portion of the Pareto front is provided due to non-uniform convergence, the decision maker may neglect some important options~\cite{weise2012evolutionary}. Having a narrow Pareto front becomes essential, in particular, when designing a device that may be required to work in different operational conditions, for example with stringent execution time constraints and less strict power dissipation ones or, on the contrary, with low power requirements and tolerance to high execution time.
